Static Race Detection in OS Kernels by Mining Locking Rules

### 2023-08-03 1.introduction.tex
Update

(T3) \sys uses a {\em pattern-based estimation} to extract harmful races that 
can trigger memory or logic bugs such as null-pointer dereferences, data 
inconsistencies and double fetches. Some data races are benign, and developers 
do not put effort into repairing them because they can not cause serious memory 
or logic bugs. Therefore, it is important to estimate the harmfulness a data 
race can bring. We observe that harmful data races can be estimated by specific 
patterns. For example, if a raced data performs as an operand of a dereference 
instruction, a null-pointer dereference can occur; if a raced data structure is 
accessed for more than once and each access gets its different fields, a data 
inconsistency can occur; if a data race is caused by an unprotected write, and 
it is likely to introduce double fetches.Based on this observation, \sys 
exploits some patterns to extract harmful data races that can cause memory or 
logic bugs automatically.

We have implemented \sys with LLVM~\cite{clang} and Z3~\cite{z3}. 
\sys performs automated static analysis on the LLVM bytecode of the 
checked OS kernel. Overall, we make three main contributions in this paper:

\begin{itemize}
	\item We analyze the challenges to detect races in kernels, and propose 
	three key techniques to address	these challenges: (T1) an {\em alias-aware 
	rule mining method} to automatically deduce locking rules; (T2) a {\em 
	lock-usage analysis} to filter out false positives caused by code that can 
	not execute concurrently; (T3) a {\em pattern-based estimation} to extract 
	harmful data races that can cause memory or logic bugs such as null-pointer 
	dereferences, data inconsistencies and double fetches.		
	\item Based on these three key techniques, we design a practical static 	
	analysis approach named \sys, to effectively detect data races and 	
	estimate the harmfulness of these data races in OS kernels.
	\item We have evaluated \sys on Linux 6.2, and find 273 real data races, 	
	with a false positive rate of 19.9\%. Among these data races, 88 are 
	estimated to be harmful. We have reported these harmful bugs to Linux 
	kernel developers, and 32 of them have been confirmed.
\end{itemize}

>>>>

(T3) \sys uses a {\em pattern-based estimation} to extract harmful races that 
can trigger memory or logic bugs such as null-pointer dereferences, data 
inconsistencies and double fetches. Some data races are benign, and developers 
do not put effort into repairing them because they can not cause serious memory 
or logic bugs. Therefore, it is important to estimate the harmfulness a data 
race can bring. We observe that harmful data races can be estimated by specific 
patterns. For example, if a raced data performs as an operand of a dereference 
instruction, a null-pointer dereference can occur; if a raced data structure is 
accessed for more than once and each access gets its different fields, a data 
inconsistency can occur; if a data race is caused by an unprotected write, it 
is likely to introduce double fetches.Based on this observation, \sys exploits 
some patterns to extract harmful data races that can cause memory or logic bugs 
automatically.

We have implemented \sys with LLVM~\cite{clang}. \sys performs automated static 
analysis on the LLVM bytecode of the checked OS kernel. Overall, we make three 
main contributions in this paper:

\begin{itemize}
	\item We analyze the challenges to detect races in kernels, and propose 
	three key techniques to address	these challenges: (T1) an {\em alias-aware 
	rule mining method} to automatically deduce locking rules; (T2) a {\em 
	lock-usage analysis} to filter out false positives caused by code that can 
	not execute concurrently; (T3) a {\em pattern-based estimation} to extract 
	harmful data races that can cause memory or logic bugs such as null-pointer 
	dereferences, data inconsistencies and double fetches.		
	\item Based on these three key techniques, we design a practical static 	
	analysis approach named \sys, to effectively detect data races and 	
	estimate the harmfulness of these data races in OS kernels.
	\item We have evaluated \sys on Linux 6.2, and find 273 real data races, 	
	with a false positive rate of 19.9\%. Among these data races, 87 are 
	estimated to be harmful. We have reported these harmful data races to Linux 
	kernel developers, and 32 of them have been confirmed.
\end{itemize}

### 2023-08-03 2.motivation.tex
Update

\PP{C1: Getting locking rules.} The relationship between variables and locks is 
not well documented in OS kernels, making it hard to determine whether a 
specific variable should be protected by a lock and which lock is required, 
even for an expert developer. And thus existing 
annotation-basedapproaches~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, 
Anderson:PLDI09, Zhou:MICRO19, Flanagan:PASTE01, Flanagan:PLDI00, 
Sadowski:PLATEAU14, ClangThreadSafety, Blackshear:OOPSLA18} are difficult to 
apply to race detection in OS kernels. Other approaches~\cite{Choi:PLDI02, 
Engler:SOSP03, Voung:FSE07, Pratikakis:PLDI06, Naik:PLDI06} employ 
lockset-based analysis to detect data races automatically, but they do not 
consider alias relationships~\cite{Voung:FSE07, Engler:SOSP03} or just use 
imprecise flow-insensitive alias analysis~\cite{Choi:PLDI02, 	
Pratikakis:PLDI06, Naik:PLDI06}. However, due to the heavy use of pointers and 
data structure fields in kernel code, the alias relationships between variables 
can be very complex, and thus lacking effective alias analysis can introduce 
both false positives and false negatives.

>>>>

\PP{C1: Getting locking rules.} The relationship between variables and locks is 
not well documented in OS kernels, making it hard to determine whether a 
specific variable should be protected by a lock and which lock is required, 
even for an expert developer. And thus existing 
annotation-based approaches~\cite{Boyapati:OOPSLA02, Anderson:PLDI08, 
Anderson:PLDI09, Zhou:MICRO19, Flanagan:PASTE01, Flanagan:PLDI00, 
Sadowski:PLATEAU14, ClangThreadSafety, Blackshear:OOPSLA18} are difficult to 
apply to race detection in OS kernels. Other approaches~\cite{Choi:PLDI02, 
Engler:SOSP03, Voung:FSE07, Pratikakis:PLDI06, Naik:PLDI06} employ 
lockset-based analysis to detect data races automatically, but they do not 
consider alias relationships~\cite{Voung:FSE07, Engler:SOSP03} or just use 
imprecise flow-insensitive alias analysis~\cite{Choi:PLDI02, 	
Pratikakis:PLDI06, Naik:PLDI06}. However, due to the heavy use of pointers and 
data structure fields in kernel code, the alias relationships between variables 
can be very complex, and thus lacking effective alias analysis can introduce 
both false positives and false negatives.

### 2023-08-03 3.technique.tex
Update

\subsection{Alias-Aware Rule Mining Method}
\label{subsec_rule_mining}
The relationship between variables and locks is not well documented in OS 
kernels, but it can be inferred from the kernel code. Specifically, a variable 
is often protected by the lock exist in the same data structure. And thus if a 
variable is accessed after acquiring a lock existing in the same data structure 
in most cases, it is likely to be protected by the lock. Whether a variable and 
the protecting lock exist in the same data structure can be determined through 
an alias graph~\cite{Li:ASPLOS22, Kastrinis:CC18}, by finding their common 
ancestor. Based on this insight, we propose an {\em alias-aware rule mining 
method} to deduce locking rules automatically. Moreover, with benefits from 
precise field-sensitive alias relationships of alias graph, our alias-aware 
rule mining method can effectively extract accessed filed and its corresponding 
lock field.

@@

Based on the alias graph, our alias-aware rule mining method performs an 
inter-procedural, path-based~\cite{Li:ASPLOS22}, field-sensitive and 
alias-aware analysis to effectively mine locking rules about whether accesses 
to a specific data structure field should be protected and which data structure 
field the protecting lock exist in. Overall, our alias-aware rule mining method 
has three main stages shown in Figure~\ref{fig_workflow}. In Stage 1, it 
extracts key fields that may be protected by a lock. In Stage 2, it collects 
all accesses to key fields as well as code paths these accesses exist in. In 
Stage 3, it deduces locking rules by calculating the proportion of field 
accesses protected by a specific lock in all field accesses.

@@

\PP{S3: Locking-rule deduction.} After collecting all accesses to key fields, 
the analysis deduces locking rules based on statistics. However, on the one 
hand, distinguishing different accesses to the same data structure field by 
different program sites is not fine-grained enough, because the access to a 
field and the acquirement to a specific lock are often packaged in a special 
function, and all accesses under different calling context are regarded as the 
same access in this strategy. On the other hand, distinguishing accesses by 
different code paths can also suffer from inaccuracy when a function contains 
many branch statements, because accesses to the same field is regarded as 
different in different code paths, causing numerous accesses to the same field 
due to large quantities of code paths. Based on this consideration, the 
analysis distinguishes accesses to the same data structure field by different 
calling contexts. Specifically, given a key field {\em f} and a lock {\em l}, 
the analysis first finds all access to {\em f} with lock {\em l} from the 
collected field accesses in Stage 2, and gets the number {\em 
locked\_access\_num} by counting different calling contexts extracted from the 
code paths of these field accesses. Then, the analysis finds all access to {\em 
f} (no matter whether a lock is held), and gets the number {\em 
all\_access\_num} in the same way as {\em locked\_access\_num}. If the ratio 
{\em locked\_access\_num} / {\em all\_access\_num} is larger than a given 
threshold, and there is at least one write access to {\em f}, the data 
structure field {\em f} is inferred to be protected by the lock {\em l}. After 
deducting locking rules, the analysis detects data races by checking whether a 
field access validates any locking rules.

@@

After extracting key fields, the method collects all accesses to key fields 
with lock-set analysis and the alias graph. Take the code path {\tt Line 14, 
Line 15, Line 5, Line 6, Line 7, Line 8, Line 9, Line 16 and Line 17} as an 
example, through a lock instruction at Line 6, the analysis records the 
acquired lock {\em Dev.lock} into the lock set. When analyzing the read 
instruction at Line 7, the analysis finds the node of the accessed field 
($\mathit{n_2}$) and the node of the lock stored in the lock set 
($\mathit{n_4}$) has the same ancestor ($\mathit{n_1}$), and thus records a 
field access <{\em Dev.flag}, {\em Dev.lock}>. Similar to the read instruction 
at Line 7, the method also records a field access <{\em Dev.flag*}, {\em 
Dev.lock}> for the write instruction at Line 8. Then through an unlock 
operation at Line 9, the analysis removes the lock {\em Dev.lock} from the lock 
set. As a result, the key fields {\em Dev.flag} and {\em Dev.flag*} are 
accessed without acquiring any lock at Lines 16 and 17, and thus the analysis 
records two field accesses <{\em Dev.flag}, {\em NULL}> and <{\em Dev.flag*}, 
{\em NULL}>. The final field accesses are shown in 
Figure~\ref{fig_demo_rule_mining}(c).

@@

\subsection{Lock-Usage Analysis}
\label{subsec_lock_usage_analysis}
Our analysis takes functions that have no caller function as the entries like 
existing work~\cite{Li:ASPLOS22}, and performs a path-based analysis starting 
from these entry functions. It assumes that all codes reaching from entry 
functions can execute concurrently, to reduce false negatives. However, this 
assumption is too conservative and can introduce many false positives because 
not all entry functions can execute concurrently in fact. We observe that each 
kernel module has an initialization phase, after which many functions can be 
called concurrently by other modules. When performing initialization, the 
kernel module serially initializes the locks and prepares other data for 
subsequent operations. And thus functions with the lock initialization and 
functions called by them tend not to execute concurrently. Based on this 
observation, we propose a lock-usage analysis to filter out false positives 
caused by code that can not execute concurrently.

@@

\noindent{\textbf{\em Example.}} Figure~\ref{fig_demo_lock_usage} shows a false 
data race filtered out by our lock-usage analysis in the Linux DRM scheduler. 
For example in {\em drm\_sched\_entity\_set\_priority()}, the accesses to {\em 
entity->priority} is protected by the lock {\em entity->rq\_lock} in most 
cases. Therefore, {\em entity-priority} is deduced to be protected by the lock 
{\em entity->rq\_lock}. But in {\em drm\_sched\_entity\_init()}, {\em 
entity->priority} is written without acquiring the lock {\em entity->rq\_lock} 
and thus introducing a possible data race. However, the lock initialization 
function {\em spin\_lock\_init()} is called at Line 86 by {\em 
drm\_sched\_entity\_init()}. And thus {\em drm\_sched\_entity\_init()} is 
inferred to execute in the module initialization phase and can not execute 
concurrently with other functions, and this is a false data race.

>>>>

\subsection{Alias-Aware Rule Mining Method}
\label{subsec_rule_mining}
The relationship between variables and locks is not well documented in OS 
kernels, but it can be inferred from the kernel code. Specifically, a variable 
is often protected by the lock exists in the same data structure. And thus if a 
variable is accessed after acquiring a lock existing in the same data structure 
in most cases, it is likely to be protected by the lock. Whether a variable and 
the protecting lock exist in the same data structure can be determined through 
an alias graph~\cite{Li:ASPLOS22, Kastrinis:CC18}, by finding their common 
ancestor. Based on this insight, we propose an {\em alias-aware rule mining 
method} to deduce locking rules automatically. Moreover, with benefits from 
precise field-sensitive alias relationships of alias graph, our alias-aware 
rule mining method can effectively extract accessed filed and its corresponding 
lock field.

@@

Based on the alias graph, our alias-aware rule mining method performs an 
inter-procedural, path-based~\cite{Li:ASPLOS22}, field-sensitive and 
alias-aware analysis to effectively mine locking rules about whether accesses 
to a specific data structure field should be protected and which data structure 
field the protecting lock exists in. Overall, our alias-aware rule mining 
method has three main stages shown in Figure~\ref{fig_workflow}. In Stage 1, it 
extracts key fields that may be protected by a lock. In Stage 2, it collects 
all accesses to key fields as well as code paths these accesses exist in. In 
Stage 3, it deduces locking rules by calculating the proportion of field 
accesses protected by a specific lock in all field accesses.

@@

\PP{S3: Locking-rule deduction.} After collecting all accesses to key fields, 
the analysis deduces locking rules based on statistics. However, on the one 
hand, distinguishing accesses to the same data structure field by different 
program sites is not fine-grained enough, because the access to a field and the 
acquisition of a specific lock are often packaged in a special function, and 
all accesses under different calling context are regarded as the same access in 
this strategy. On the other hand, distinguishing accesses by different code 
paths can also suffer from inaccuracy when a function contains many branch 
statements, because accesses to the same field is regarded as different in 
different code paths, causing numerous accesses to the same field due to large 
quantities of code paths. Based on this consideration, the analysis 
distinguishes accesses to the same data structure field by different calling 
contexts. Specifically, given a key field {\em f} and a lock {\em l}, the 
analysis first finds all access to {\em f} with lock {\em l} from the collected 
field accesses in Stage 2, and gets the number {\em locked\_access\_num} by 
counting different calling contexts extracted from the code paths of these 
field accesses. Then, the analysis finds all access to {\em f} (no matter 
whether a lock is held), and gets the number {\em all\_access\_num} in the same 
way as {\em locked\_access\_num}. If the ratio {\em locked\_access\_num} / {\em 
all\_access\_num} is larger than a given threshold, and there is at least one 
write access to {\em f}, the data structure field {\em f} is inferred to be 
protected by the lock {\em l}. After deducting locking rules, the analysis 
detects data races by checking whether a field access validates any locking 
rules.

@@

After extracting key fields, the method collects all accesses to key fields 
with lock-set analysis and the alias graph. Take the code path {\tt Line 14, 
Line 15, Line 5, Line 6, Line 7, Line 8, Line 9, Line 16 and Line 17} as an 
example, through a lock instruction at Line 6, the analysis records the 
acquired lock {\em Dev.lock} into the lock set. When analyzing the read 
instruction at Line 7, the analysis finds the node of the accessed field 
($\mathit{n_2}$) and the node of the lock stored in the lock set 
($\mathit{n_4}$) has the same ancestor ($\mathit{n_1}$), and thus records a 
field access <{\em Dev.flag}, {\em Dev.lock}, read>. Similar to the read 
instruction at Line 7, the method also records a field access <{\em Dev.flag*}, 
{\em Dev.lock}, write> for the write instruction at Line 8. Then through an 
unlock operation at Line 9, the analysis removes the lock {\em Dev.lock} from 
the lock set. As a result, the key fields {\em Dev.flag} and {\em Dev.flag*} 
are accessed without acquiring any lock at Lines 16 and 17, and thus the 
analysis records two field accesses <{\em Dev.flag}, {\em NULL}, read> and 
<{\em Dev.flag*}, {\em NULL}, write>. The final field accesses are shown in 
Figure~\ref{fig_demo_rule_mining}(c).

@@

\subsection{Lock-Usage Analysis}
\label{subsec_lock_usage_analysis}
Our analysis takes functions that have no caller function as the entries like 
existing work~\cite{Li:ASPLOS22}, and performs a path-based analysis starting 
from these entry functions. It assumes that all codes reaching from entry 
functions can execute concurrently, to reduce false negatives. However, this 
assumption is too conservative and can inevitably introduce false positives 
because not all entry functions can execute concurrently in fact. We observe 
that each kernel module has an initialization phase, after which many functions 
can be called concurrently by other modules. When performing initialization, 
the kernel module serially initializes the locks and prepares other data for 
subsequent operations. And thus functions with the lock initialization and 
functions called by them tend not to execute concurrently. Based on this 
observation, we propose a lock-usage analysis to filter out false positives 
caused by code that can not execute concurrently.

@@

\noindent{\textbf{\em Example.}} Figure~\ref{fig_demo_lock_usage} shows a false 
data race filtered out by our lock-usage analysis in the Linux DRM scheduler. 
For example in {\em drm\_sched\_entity\_set\_priority()}, the access to {\em 
entity->priority} is protected by the lock {\em entity->rq\_lock} in most 
cases. Therefore, {\em entity-priority} is deduced to be protected by the lock 
{\em entity->rq\_lock}. But in {\em drm\_sched\_entity\_init()}, {\em 
entity->priority} is written without acquiring the lock {\em entity->rq\_lock} 
and thus introducing a possible data race. However, the lock initialization 
function {\em spin\_lock\_init()} is called at Line 86 by {\em 
drm\_sched\_entity\_init()}. And thus {\em drm\_sched\_entity\_init()} is 
inferred to execute in the module initialization phase and can not execute 
concurrently with other functions, and this is a false data race.

### 2023-08-03 4.framework.tex
Update

Based on the three key techniques in Section~\ref{sec_technique}, we develop a 
practical static approach named \sys, to detect data races in OS kernels, and 
estimate the harmfulness of these data races. We have implemented \sys with 
Clang~\cite{clang} and Z3~\cite{z3}. \sys can automatically mine locking rules, 
detects data races and estimates the harmfulness of detected data races. 
Figure~\ref{fig_architecture} shows the architecture of \sys, which has four 
phases:

@@

\PP{P3: Data race detection and false-positive filtering.} The data race 
detector detects data races by checking whether a given access violates the 
rules mined by our locking-rule miner. After finding a possible data race, the 
data race filter uses our locking-usage analysis to determine whether it is a 
false positive. Besides, for a given data race, there may be multiple code 
paths from the entry function to its problematic instruction, and thus many 
repeated data races can be reported. To drop repeated data races, for a new 
possible data race, the data race filter checks whether its problematic 
instruction is identical to any existing data race. If so, this possible data 
race is regarded as repeated and dropped.

\PP{P4: Harmfulness estimation.} The harmfulness estimator uses our 
pattern-based estimation to estimate the harmfulness of detected data races, 
and reports data races that can cause memory or logic bugs. With these results, 
developers can focus on those harmful data races.

>>>>

Based on the three key techniques in Section~\ref{sec_technique}, we develop a 
practical static approach named \sys, to detect data races in OS kernels and 
estimates the harmfulness of detected data races automatically. We have 
implemented \sys with Clang~\cite{clang}. Figure~\ref{fig_architecture} shows 
the architecture of \sys, which has four phases:

@@

\PP{P3: Data race detection and false-positive filtering.} The data race 
detector detects data races by checking whether a given access violates the 
rules mined by our locking-rule miner. After finding a possible data race, the 
data race filter uses our locking-usage analysis to check whether it is a 
false positive. Besides, for a given data race, there may be multiple code 
paths from the entry function to its problematic instruction, and thus many 
repeated data races can be reported. To drop repeated data races, for a new 
possible data race, the filter checks whether its problematic 
instruction is identical to any existing data race. If so, this possible data 
race is dropped.

\PP{P4: Harmfulness estimation.} The harmfulness estimator uses our 
pattern-based estimation to estimate the harmfulness of detected data races, 
and reports data races that can cause memory or logic bugs. With these results, 
developers can focus on data races that can bring harmful impacts.

### 2023-08-03 5.evaluation.tex
Update

\begin{table}[tbph]
	\tablecaption{Detection results of Linux 6.2.}
	\label{tbl_bug_detection}
	\renewcommand{\arraystretch}{1}
	\setlength\tabcolsep{2pt}
	\noindent{\scriptsize
		\begin{center}
			\begin{tabular}{p{1.5cm}|l|c}
				\hline
				\multicolumn{2}{c|}{\textbf{Description}} & \textbf{\sys}  
				\\ \hline
				\multirow{2}{1.5cm}{\textbf{{\em Code analysis}}} & 
				Source files (analyzed/all) & xxxK/xxxK
				\\ \cline{2-3}
				& Source code lines (analyzed/all) & xxxM/xxxM 
				\\ \cline{2-3}
				\hline
				\multirow{2}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
				Key fields / total variables& 
				\\ \cline{2-3}
				& Mined locking rules &
				\\ \cline{2-3}
				\hline
				\multirow{2}{1.5cm}{\textbf{{\em Data race detection}}} & 
				Detected data races (real / all) & 273 / 341
				\\ \cline{2-3}
				& Dropped data races by lock-usage analysis & 63
				\\ \cline{2-3}
				\hline
				\multirow{4}{1.5cm}{\textbf{{\em Harmfulness estimation}}}
				& Null-pointer dereference (confirmed / all) & 15 / 20
				\\ \cline{2-3}
				& Data inconsistency (confirmed / all) & 7 / 10
				\\ \cline{2-3}
				& Double fetch (confirmed / all) & 10 / 57
				\\ \cline{2-3}
				& Total harmful data races (confirmed / all) & 32 / 88
				\\ \cline{2-3}
				\hline
				\multirow{3}{1.5cm}{\textbf{{\em Time usage}}} & 
				Key-field extraction & 
				\\ \cline{2-3}
				& Data-race detection &
				\\ \cline{2-3}
				& Total time &
				\\ \cline{2-3}
				\hline
			\end{tabular}
	\end{center}}
\end{table}

@@

\PP{Code analysis.} \sys can scale to large code bases of OS kernels, and it in 
total analyzes xxxM lines of code in xxxK source code files within xxx hours. 
The remaining xxxM lines of code in xxxK source files are not analyzed, as they 
are not enabled by the {\em allyesconfig} for the x86-64 architecture. We 
believe that \sys can also find more data races in other architectures with 
proper configuration.

@@

\PP{Data race estimation.} Many data races are benign and can not cause memory 
or logic bugs, and thus developers are unwilling to put effort into repairing 
them. We exploit three patterns to detect null-pointer dereferences, data 
inconsistencies and double fetches as introduced in 
Section~\ref{subsec_estimation}, and find 88 data races in these patterns. We 
report them to developers and 32 of them have been confirmed and fixed by them. 
We still wait for the response of other data races. Moreover, one of the 
developers wonders if \sys can be used in their CI to detect these problems. 
The results show that our pattern-based estimation can extract harmful data 
races effectively, and can considerably reduce the workload of developers.

\subsection{False Positives and Negatives}
\label{subsec_false_pos_neg}
\PP{False positives.} \sys reports 68 false data races, and through manually 
checking these false data races, we find that they are introduced for three 
main reasons:

@@

Finally, \sys exploits a lock-usage analysis to filter out false data races 
caused by code that can not execute concurrently. Specifically, the lock-usage 
analysis extracts all functions that are reachable from the lock initialization 
functions in a call graph with the union-find set, and assumes that these 
functions can not execute concurrently. However, some functions can be called 
at multiple program sites, and some of the program sites may be not reachable 
from the lock initialization functions, but data races that exist in these 
functions are all dropped by our lock-usage analysis.

@@

\PP{Null-pointer dereference in the HDA sound driver.} In 
Figure~\ref{fig_case_bugs}(a), the functions {\em snd\_hdac\_regmap\_exit()} 
and {\em snd\_hdac\_regmap\_sync()} can execute concurrently. In Thread 2, the 
variable {\em codec->regmap} is checked by an if statement in the function {\em 
snd\_hdac\_regmap\_sync()}. Right after the condition of the if statement is 
calculated to be true, the variable {\em codec->regmap} is set to be NULL by 
the function {\em snd\_hdac\_regmap\_exit()} in Thread 1. And then the function 
{\em regcache\_sync()} is called in Thread 2, with the argument {\em 
codec->regmap}, after acquiring the lock {\em codec->regmap\_lock}. In the 
called function, the variable {\em codec->regmap} is dereferenced through 
{\em map->lock()}. In this execution case, the data structure field {\em 
codec->regmap} is first assigned with NULL in Thread 1, and then dereferenced 
in Thread 2, and thus cause a null-pointer dereference.

@@

\PP{Double fetch in the LPFC SCSI driver.} In Figure~\ref{fig_case_bugs}(c), 
the value of the shared variable {\em phba->fcf.fcf\_flag} is set to 0 by the 
function {\em lpfc\_unregister\_fcf\_rescan()} in Thread 2, without acquiring 
the lock {\em phba->hbalock}, and this can cause double fetches in several 
program sites. For example, if {\em phba->fcf.fcf\_flag} is set to 0 by Thread 
2 between the two accesses to it in the function {\em lpfc\_register\_fcf()} in 
Thread 1 (Lines 1873 and 1874), the values of the two accesses can be 
different, causing unpredictable behaviors.

>>>>

\begin{table}[tbph]
	\tablecaption{Detection results of Linux 6.2.}
	\label{tbl_bug_detection}
	\renewcommand{\arraystretch}{1}
	\setlength\tabcolsep{2pt}
	\noindent{\scriptsize
		\begin{center}
			\begin{tabular}{p{1.5cm}|l|c}
				\hline
				\multicolumn{2}{c|}{\textbf{Description}} & \textbf{\sys}  
				\\ \hline
				\multirow{2}{1.5cm}{\textbf{{\em Code analysis}}} & 
				Source files (analyzed/all) & xxxK/xxxK
				\\ \cline{2-3}
				& Source code lines (analyzed/all) & xxxM/xxxM 
				\\ \cline{2-3}
				\hline
				\multirow{2}{1.5cm}{\textbf{{\em Locking-rule mining}}} & 
				Variables in key fields / total variables& 
				\\ \cline{2-3}
				& Mined locking rules &
				\\ \cline{2-3}
				\hline
				\multirow{2}{1.5cm}{\textbf{{\em Data race detection}}} & 
				Detected data races (real / all) & 273 / 341
				\\ \cline{2-3}
				& Dropped data races by lock-usage analysis & 63
				\\ \cline{2-3}
				\hline
				\multirow{4}{1.5cm}{\textbf{{\em Harmfulness estimation}}}
				& Null-pointer dereference (confirmed / all) & 15 / 20
				\\ \cline{2-3}
				& Data inconsistency (confirmed / all) & 7 / 10
				\\ \cline{2-3}
				& Double fetch (confirmed / all) & 10 / 57
				\\ \cline{2-3}
				& Total harmful data races (confirmed / all) & 32 / 87
				\\ \cline{2-3}
				\hline
				\multirow{3}{1.5cm}{\textbf{{\em Time usage}}} & 
				Key-field extraction & 
				\\ \cline{2-3}
				& Data-race detection &
				\\ \cline{2-3}
				& Total time &
				\\ \cline{2-3}
				\hline
			\end{tabular}
	\end{center}}
\end{table}

@@

\PP{Code analysis.} \sys can scale to large code bases of OS kernels, and it in 
total analyzes xxxM lines of code in xxxK source code files within xxx hours. 
The remaining xxxM lines of code in xxxK source files are not analyzed, as they 
are not enabled by the {\em allyesconfig} for the x86-64 architecture.

@@

\PP{Data race estimation.} Many data races are benign and can not cause memory 
or logic bugs, and thus developers are unwilling to put effort into repairing 
them. We exploit three patterns to detect null-pointer dereferences, data 
inconsistencies and double fetches as introduced in 
Section~\ref{subsec_estimation}, and find 87 data races in these patterns. We 
report them to developers and 32 of them have been confirmed and fixed by them. 
We still wait for the response of other data races. Moreover, one of the 
developers wonders if \sys can be used in their CI to detect these problems. 
The results show that our pattern-based estimation can extract harmful data 
races effectively, and can considerably reduce the workload of developers.

\subsection{False Positives and Negatives}
\label{subsec_false_pos_neg}
\PP{False positives.} \sys reports 68 false data races, and through manually 
checking these false data races, we find that they are introduced for two main 
reasons:

@@

Finally, \sys exploits a lock-usage analysis to filter out false data races 
caused by code that can not execute concurrently. Specifically, the lock-usage 
analysis extracts all functions that are reachable from the lock initialization 
functions in a call graph with the union-find set, and assumes that these 
functions can not execute concurrently. However, A function can be called under 
different calling contexts, and some of the calling contexts do not contain 
lock initialization operations, but data races that exist in these functions 
are all dropped by our lock-usage analysis, if any calling context contains a 
lock initialization operation.

@@

\PP{Null-pointer dereference in the HDA sound driver.} In 
Figure~\ref{fig_case_bugs}(a), the functions {\em snd\_hdac\_regmap\_exit()} 
and {\em snd\_hdac\_regmap\_sync()} can execute concurrently. In Thread 2, the 
variable {\em codec->regmap} is checked by an if statement in the function {\em 
snd\_hdac\_regmap\_sync()}. Right after the condition of the if statement is 
calculated to be true, the variable {\em codec->regmap} is set to NULL by the 
function {\em snd\_hdac\_regmap\_exit()} in Thread 1. And then the function 
{\em regcache\_sync()} is called in Thread 2, with the argument {\em 
codec->regmap}, after acquiring the lock {\em codec->regmap\_lock}. In the 
called function, the variable {\em codec->regmap} is dereferenced through 
{\em map->lock()}. In this execution case, the data structure field {\em 
codec->regmap} is first assigned with NULL in Thread 1, and then dereferenced 
in Thread 2, and thus cause a null-pointer dereference.

@@

\PP{Double fetch in the LPFC SCSI driver.} In Figure~\ref{fig_case_bugs}(c), 
the value of the shared variable {\em phba->fcf.fcf\_flag} is set to 0 by the 
function {\em lpfc\_unregister\_fcf\_rescan()} in Thread 2, without acquiring 
the lock {\em phba->hbalock}, and this can cause double fetches in several 
program sites, even if other accesses are all protected by a lock. For example, 
if {\em phba->fcf.fcf\_flag} is set to 0 by Thread 2 between the two accesses 
to it in the function {\em lpfc\_register\_fcf()} in Thread 1 (Lines 1873 and 
1874), the values of the two accesses can be different, causing unpredictable 
behaviors.
